%Thesis: Multi-dimensional spaces are better visualized through slice based views

We live in a three-dimensional world.  Ourselves and what we can interact with
are in three dimensions.  We learn about our world by studying the various
phenomena around us.  These phenomena are described as continuous processes.
In the beginning of our  education we study function plots in high school.
These give an intuitive view of one-dimensional phenomena.  By exploring the
relationship between an input factor
and output,
we can build an understanding on the relationship between the two.  We can also
compare one function plot to another. Visual inspection of these plots allows
us to see common patterns. We use our pattern recognition ability to quickly
categorize these different plots into different types of function behavior.
Function plots can also be used to describe two-dimensional phenomena. These
show the effect due to two input factors. In this case we can use the third
dimension or color encoding to show the function value.  From these plots we
can also make general statements about the ``shapes'' of the behavior like how
``peaky'' the function is or if it is monotonically increasing. These shapes
give us intuition into the underlying processes and help us learn about the
world~\cite{Palmer:1999}.

We also interact with many phenomena around us that are essentially
multi-dimensional in nature. For example, the weather in a certain location is
determined by the temperature, pressure, humidity, dew point, wind velocity,
and wind direction, among others. A change in any of the factors results in a
change in the weather. Each of these factors can be given a ``spatial
embedding.'' Then, they can be viewed as a dimension of some space.  By
``walking'' or navigating through this space we can observe the effect on the
weather due to changes in these parameters. 

Understanding multi-dimensional continuous spaces is difficult. As
three-dimensional beings we have real-world analogs for measurement, angle, and
position in three dimensions. We do not have these once we move beyond three
dimensions though. Nevertheless, visual analysis of these multi-dimensional
spaces has produced insights about the underlying
behavior~\cite{Sedlmair:2014,Gleicher:2016}. The issue is how to show more than
three dimensions on a two-dimensional screen. 

One strategy is to discretize the dataset through sampling and then use the
wealth of discrete visualization tools available. Much of the work on
multi-dimensional data analysis developed from analysis of 
tabular data. These datasets are often recorded from real-world
events such as census, species, or text data and contain many different aspects
about each entity. Thus, these datasets are inherently multi- or high-dimensional.
Each different aspect of the data items creates a dimension to be analyzed. 
In the case of these data the mental model is that of discrete objects
like humans, plants, or documents. Since the mental model is discrete in this
case it makes sense to use data processing and visualization tools designed for
discrete data.  However, our mental model for physical data is a continuous 
one. Therefore, the
discrete data paradigm breaks our mental model~\cite{Tory:2004a,Liu:2010a}. 
Breaking the mental model means that the visualization is not conveying the
complexities of the continuous phenomena.
Rather, we
should use visualization techniques purpose-built for continuous data.

While not as
extensively developed, there is previous work on visualizing multi-dimensional,
continuous data. These techniques can be broadly classified into projection,
topological, and slicing methods.  
Projection methods attempt to
distort the multi-dimensional object in order to view it on a two-dimensional
screen. With projection techniques we can preserve distance, direction, size,
or angles, but not all of 
these~\cite{Snyder:1987}. 
Depending on the
projection method, we may see radically different representations.  The issue
is that it is not clear from the resulting visualization what sort of
transformation was performed on the data.  Thus it can be difficult to
reconstruct the mental model of the multi-dimensional object.  One of the most
often seen multi-dimensional projection techniques is the Schlegel
diagram~\cite{Sommerville:1929} which picks a ``face'' of a polytope and projects the
remaining faces inside it. Thus, this technique only works for 4D polytopes.
Topological methods search the continuous dataset for values of interest, such
as critical points or contours.  Topological visualization techniques also
suffer from the issue of unclear transformation. It is difficult to relate the
resulting visualization back to features in the multi-dimensional object.

Slice-based views of multi-dimensional continuous spaces have not been explored
as extensively as other options.  This work began with the advent of
HyperSlice~\cite{Wijk:1993}.  HyperSlice extends the idea of slicing from
medical imaging to any number of dimensions.  HyperSlice provides the framework
for visualizing multi-dimensional continuous objects as a set of
two-dimensional slices.  There are $d \choose 2$ subpanels, one for each pair
of dimensions. Each panel shows a 2D slice of the object. The horizontal axis
shows one dimension and the vertical axis shows another dimension. Essentially,
each sub-plot of HyperSlice shows a 2D function plot. With 2D slices of solid
multi-dimensional objects color is often used to encode value. 

My work is inspired by the HyperSlice technique. Van Wijk and van Liere
introduced the idea of using slice-based views of multi-dimensional data.
However, they did not expand on what data types and tasks are involved in
multi-dimensional continuous data analysis.  I build on their work,
investigating the usefulness of slice-based views of continuous
multi-dimensional datasets. I also identified tasks involved in
multi-dimensional data analysis. The task analysis informed the development of
one- and two-dimensional slice-based views.

In this thesis I will explore
the possibilities of these slice-based views. Through a number of case
study examples, I will demonstrate the power of these views and ways to
address their shortcomings.

\subsection{Multi-dimensional spaces}
\label{sec:motivation:multi-d}

There are a number of domains where one can apply the analysis of continuous
multi-dimensional data.  As of yet, there has not been a comprehensive data and
task analysis for multi-dimensional continuous data analysis. For discrete
data, there are several task 
analyses~\cite{Shneiderman:1996,Brehmer:2013,Amar:2004}. 
However, they are
focused on identifying and selecting particular data items. Continuous datasets
consist of ranges of values as well as functions. Functions can be seen as a
mapping from ranges of numbers to other ranges. The analysis task here is to
study these ranges, their relationships to each other, and the mappings between
them. Tasks addressing these
have not yet been covered by visualization task analyses. Thus,
there is no comprehensive source for what analysis tasks one wants to perform
given a continuous multi-dimensional dataset.  Work in this area has
traditionally focused on developing a specific visualization for a specific
task. For example, topological spines extracts critical points from a scalar
field~\cite{Correa:2011}. One goal of this thesis is to develop this task
taxonomy for visualization of continuous multi-dimensional data.


These domains can be broadly classified into two types based on their analysis
tasks. One type, \emph{manifold} analysis, deals with understanding the
relationship between inputs and outputs. This is a functional relationship.
The user wants to inspect how changes in the inputs (independent variables)
affect the outputs (dependent variables). One can also perform \emph{shape}
analysis. Here, in terms of the analysis tasks, there is no identification of
independent and dependent variables. We look at each of these two in turn.

\subsection{Manifolds}
\label{sec:manifolds}

Studying the mapping between continuous ranges means studying functional
relationships and thus manifolds.  The critical issue is understanding the
relationship between independent and dependent variables.  Subtasks in manifold
analysis include examining critical points, assessing the sensitivity of
parameters, and understanding the shape of the manifold.  
One area where
understanding the manifold is important is analyzing optimization surfaces and
functions.  In this case, the identification of extrema is important for
understanding how many and the relative location of local optima. In addition,
we want to understand the degree to which these are extrema. These can result
in global optimization algorithms ``getting stuck'' in local optima rather than
continuing to search for the global optimum. Optimization algorithms need to be
carefully tuned to properly detect these features and ignore them where
necessary.

Simulations can be used to run experiments that are impractical or impossible
in the real world.  Simulation analysis is another area where the analysis
tasks, in the abstract, are examining functions. If we look back at the weather
simulation from before, the inputs to the function are things like the
temperature and pressure.  The output is, for example, the likelihood of rain
the next day. The function is the simulation itself. Computer simulations are
deterministic.  A deterministic simulation has a fixed mapping from each unique
input parameter configuration to an output value. This is the same as a
functional relationship. The sensitivity and extrema are also important to
simulation analysis.  Thus, these can all be analyzed with visualizations of a
manifold.

To date, manifold visualizations have concentrated on a particular analysis
task or a particular application domain. For example, visualizations of the
Morse-Smale complex~\cite{Gerber:2010} are focused on showing only critical
points of the manifold. As with any visualization designed for a specific task,
they must be used in combination with other views for visual analysis of
domain-specific data. Domain-specific visualizations often used linked views to
show different aspects of data to accomplish multiple tasks at once. However,
they are purpose built for a specific domain. While techniques may transfer
from one domain to another~\cite{Sedlmair:2012}, it is not always clear how.
My goal is to unify these methods to a certain extent.  As I will show in
\autoref{chp:sliceplorer}, slice-based views of manifolds can be used for a
wide variety of tasks in a wide number of domains.

\subsection{Shapes}
\label{sec:shapes}

One may also want to understand the relationship or correlation between
multiple continuous values. In the manifold analysis case we have the notion
of independent and dependent variables. This classification does not exist 
here.
In this case
we want to study the relationship of all variables. Careful study of the shape
of the dataset can give insight into the relationship between the various
ranges of dimensions of the object. For example, one may want to know if the
overall shape is a sphere, donut, or box. In addition one may be interested in
any kinks or cusps in the dataset. Changes in the gradient and curvature of
the shape are also of great interest. These indicate changes in correlation or
relation.

The analysis tasks of multi-dimensional shapes can be applied to a number of
different areas. The study of polytopes is one such area and perhaps the most
direct application of understanding multi-dimensional continuous shapes.
Polytopes are the multi-dimensional generalization of polyhedra and polygons.
The tasks are to understand the symmetries and patterns making up the
polytopes~\cite{Ziegler:2012}. Perhaps a less obvious connection is the
analysis of the tradeoff curves in multi-objective optimization. Since we are
performing optimization we are interested in the tradeoffs amongst all the
non-dominated points~\cite{Kung:1975}. This is also known as the Pareto
front. In this case, the user wants to understand what are the costs of
reducing one or more parameters in order to increase the value of others. Cusps
or large changes in curvature in these datasets are important since they show
critical changes in the rate of tradeoff.  With a proper view of a
multi-dimensional object we can also view differences between two objects
directly. 

These two different data types and sets of tasks require different visualization
considerations. Proper visualization for manifold analysis should focus on
the relationships between independent and dependent variables. Visualizations
of shapes do not have this mapping requirement and instead focus on the 
relationships between dimensions. With this categorization in mind, we can now
examine the available visualization techniques to examine these.


\subsection{Visualizing multi-dimensional continuous spaces}
\label{sec:multi-d-challenges}

Understanding multi-dimensional space is difficult. As humans, we simply do not
have the spatial analogs in more than three dimensions. A number of methods
have been developed to extract specific features from the multi-dimensional
object. For example, when studying polytopes, the number of faces and
symmetries is very important~\cite{Ziegler:2012}. However, these only produce
an answer without sufficient context. They do not necessarily give any
intuition as to how to transfer our three-dimensional knowledge to
multi-dimensional spaces.

Visualizations of multi-dimensional spaces on a 2D screen must contend with
some sort of reduction of the information. A proper visualization must select
visual encodings that highlight the information we want to see. Any sort of
data reduction requires trade-offs. The best visualization choices acknowledge
any deficiencies to a particular visual encoding. By acknowledging these
deficiencies, we can design tools to compensate for their shortcomings while
still maintaining their advantages. Therefore, it is worth first looking at the
possible mappings of data to visual elements. Then, I present commonly used
visual encodings of multi-dimensional continuous data using these mappings.

\subsection{Encoding multi-dimensional data}

Multi-dimensional continuous data consists of a set of continuous ranges, one
for each dimension. In the case of manifold analysis, each of these ranges can
be additionally classified as ``dependent'' or ``independent'' depending on
which side of the mapping they are on. Typical visualization practice is to
give each dimension a separate visual channel. There are a number of possible
visual channels that have been identified.  The ranking of effectiveness of
visual channels (shown in \autoref{tbl:visual_encodings}) was proposed by
Bertin~\cite{Bertin:1967} and confirmed through experiments by Cleveland and
McGill~\cite{Cleveland:1984}, Mackinlay~\cite{Mackinlay:1986}, and Heer and
Bostock~\cite{Heer:2010}.  Munzner~\cite{Munzner:2014} provides a summary of
the results. We are also limited in how many channels we can use
simultaneously. According to Ware~\cite{Ware:2004}, certain channels, such as
red and green are not visually separable. 

\begin{table}
  \caption{Rankings of visual encodings of quantitative data}
  \label{tbl:visual_encodings}
  \begin{adjustbox}{max width=\linewidth}
  \begin{tabular}{llll}
    Bertin~\cite{Bertin:1967} & Cleveland and McGill~\cite{Cleveland:1984} & Mackinlay~\cite{Mackinlay:1986} & Munzner~\cite{Munzner:2014} \\
    \hline \\
     Position & Position along a common scale & Position & Position on common scale \\
     Size & Position along identical, nonaligned scales & Length & Position on unaligned scale \\
     (Grey) Value & Length & Angle & Length (1D size) \\
     Texture & Direction & Slope & Tilt/angle \\
     Color & Angle & Area & Area (2D size) \\
     Orientation & Area & Volume & Depth (3D position) \\
     Shape & Volume & Density & Color luminance\\
     & Curvature & Color saturation & Color saturation \\
     & Densities & Color hue & Curvature \\
     & Shading & & Volume (3D size) \\
     & Color saturation &           & 
  \end{tabular}
  \end{adjustbox}
\end{table}

The difficulty of visualizing a continuous multi-dimensional space on a
two-dimensional screen brings a number of challenges. We treat each dimension
separately, thus, we need several different visual channels. However, there are
simply not enough visual channels available to draw a 15-dimensional object in
a single view. This is further complicated by the fact that separate visual
channels are not necessarily visually separable.  Furthermore, each dimension
of the multi-dimensional object under study is treated equally. For example, no
particular axis of a polytope is more important than any other.  We should
encode each dimension using equally weighted effectiveness channels.  With
fewer channels available than data dimensions we either need to reduce the data
or use multiple views to properly visualize the data.


\subsection{Methods}

The common taxonomy of how to view multi-dimensional data on screen is based on
discrete data analysis. There, there are two categories: dimension reduction or
projection. With continuous data, though there is a third possibility, that of
slicing. Therefore, I view the taxonomy of \emph{continuous} multi-dimensional
data analysis methods into two categories. Data-driven methods include both
projection and dimension reduction and reduce the dimensionality of the data
before visualization. View-based methods reduce the data during the
visualization. Slicing is a view-based method.

Purely data-driven methods are commonly known as feature selection or dimension
reduction. The goal is to find a subset of dimensions that are critical to
understanding the dataset. Topological techniques take this a step further.
They discard all spatial information about the dataset and only concentrate on
the difference in function value, as in the Morse-Smale
complex~\cite{Gyulassy:2012a}, or evolution of contours, as in the contour
tree~\cite{Carr:2003a}.  Projections also synthesize the dataset into new
dimensions to show using visual channels. Principal component
analysis~\cite{Holbrey:2006} is a popular choice in this area. This rotates the
space and thus produces new set of dimensions that are a linear combination of
the input dimensions. Even this relatively simple operation (a rotation) can be
difficult to understand. For example, iPCA~\cite{Jeong:2009a} was a tool to
help users understand the effects of the dimensional transformation.

View-based methods try and produce multiple linked views of a multi-dimensional
dataset from different angles. Each view shows a subset of the dimensions.
This way we can use a proper set of visual channels for each view.  We use
interaction to link these different views. These multiple, coordinated, linked
views have been one of the biggest success stories from the visualization
community~\cite{Rao:1994}. The traditional HyperSlice~\cite{Wijk:1993}
technique falls into this category. Each panel of the HyperSlice view shows two
of the input dimensions and the value is encoded with color.
The views are linked through the focus point selection. Changing the focus point
in one sub-plot updates the other sub-plots.

My work focuses on the exploration of the combination of data-driven and
view-based methods.  Data-driven methods reduce the data in a way that we can
get a global overview of the dataset. View-based methods are much more detailed
but can only produce a local view of the data. By combining these methods I can
achieve both a global overview as well as an on-demand local view of the
dataset in a single visualization. 


\subsection{Slices}
\label{sec:slicing-advantages}

Slicing offers a number of advantages over other multi-dimensional
visualization techniques. Slicing is a direct visualization of the
multi-dimensional object. In contrast to methods like projection or dimension
reduction, slicing does not distort the dimensions in order to display them on
a two-dimensional screen. Since there is no distortion, distances in the visual
representation are directly proportional to distances in the object. This is
one of the reasons that slicing is popular in the medical imaging community.
Sizes of organs or tumors can be measured visually on screen. Additionally,
relative sizes correspond to what a doctor would expect to see in the body.
Multi-dimensional data is abstract. It lacks the correspondence to real-world
objects that the medical community uses to understand their two- and
three-dimensional datasets. Nevertheless, it is still important in
multi-dimensional data analysis to properly estimate distances and relative
sizes. 

Another benefit of the direct visualization is that users do not require
extensive training to understand the visualization. The concept of slicing
through a three-dimensional object is a familiar one. Humans are used to this
even from slicing fruits and vegetables with a knife. This concept of slicing
can be extended from this well-known metaphor to cover multiple slices of
multi-dimensional objects.

Slice views use the horizontal and vertical axes for showing the effects due to
the input parameters. These axes are the most perceptually
uniform~\cite{Stevens:1957} and are considered the most effective
(\autoref{tbl:visual_encodings}). One- or two-dimensional
plots are replicated for each combination of dimensions in order to show more
than two dimensions at once. This promotes familiarity of the visualization.
Once the user has learned to read a single panel, they can apply this knowledge
to the remaining panels. This approach follows the principal of small
multiples~\cite{Archambault:2011}. 

In order to produce a slice plot one needs to first pick a particular
\emph{focus point} in the multi-dimensional space. This focus point determines
which slices are being viewed. Selecting a good focus point a-priori is
difficult. 
It either requires a great deal of luck or careful analysis of the
dataset. This is not always possible. Slice-based views require some sort of
interactive focus point selection. Interactively browsing through the slices
requires interaction controls to give the user control over the focus point.
Furthermore, we need some kind of navigation map to show which focus points the
user has selected so that they do not become lost. Neither of these navigation
aids are well developed at this point. This need for interaction is likely one
of the reasons that slice-based views have not developed as much as projection
or topological techniques. Static views are much easier to include in papers
and don't require explanation prior to use.

The other implementation issue of slice-based views is ensuring that the
visualization can remain interactive. In this case, interactive is defined as
the rate at which the user can maintain their
concentration~\cite{Shneiderman:1987}. This is often defined at 10 frames per
second. It can be difficult to compute a 2D slice of an arbitrary complex
multi-dimensional object. 


\subsection{Upcoming}
\label{sec:thesis_outline}

My goal is to highlight the advantages of slice-based methods for
multi-dimensional data analysis. At the same time I want to address the
limitations. The end goal is to bring slice-based views into the standard
toolbox of visualizations.

In \autoref{chp:sliceplorer}, I examine how to visualize manifolds using
slices. I present Sliceplorer which is a system to view one-dimensional slices
of multi-dimensional manifolds. I use projections of these one-dimensional
slices instead of showing one focus point at a time. I also go into detail
about the tasks one wants to perform with manifold analysis.

I extend the idea of projections of slices to the second dimension in
\autoref{chp:hypersliceplorer}.  I show how this can be used to effectively
visually analyze the shape of multi-dimensional data. In many cases, this data
is given as a simplical mesh. I introduce an algorithm to compute 2D slices of
this mesh.  Using this algorithm, I show how we can visually understand
datasets like Pareto fronts or polytopes.

Finally, in \autoref{chp:rendering}, I discuss how we can take advantage of the
multi-dimensional geometry and GPU architecture to allow interactive-speed
browsing of the focus points. In addition to this algorithm, I develop a method
that can predict the amount of time needed per element to draw one frame of the
visualization. I then show how this estimation formula can be calibrated to a
particular user's hardware.


