section{Motivation}\label{motivation}

The visual analysis of multi-dimensional scalar functions is a fundamental aspect in many areas, from computational sciences (based on computer simulations) to data sciences (based on machine learning techniques) and general optimization algorithms. 
Neural networks, for instance, have shown to produce very good results that come in the form of highly non-linear response manifolds.
While understanding these manifolds would greatly help to verify
the resulting models, there
is currently no established way to inspect these multi-dimensional manifolds. %With optimization algorithms, it is important to examine the
%multi-dimensional response surface 
In case of an optimization algorithm, a good visualization of response manifolds could, for instance, help to see where possible
``holes'' (local extrema) are that the algorithm may get trapped in and fail to find the
global optimum. By visually examining the optimization functions,
one can develop new insights on how to improve the optimization
algorithm. This work is based on the assumption that visual analyses of these functions will help to increase
understanding of what they are doing, 
%(or improving comprehensibility 
as argued by Gleicher~\cite{Gleicher:2016}.

To examine these functions visually they must be reduced to be
displayed on a 2D screen. 
Currently, the two major approaches used to visualize these spaces are either
two-dimensional slices, a technique known as HyperSlice~\cite{Wijk:1993} or dimension transformation techniques. This includes dimensionality
reduction and topological methods~\cite{Correa:2011,Carr:2003a}.  HyperSlice
shows two-dimensional slices (using either a heatmap or contour view)
of the function directly around a particular focus point. It clearly shows the
behavior of the function with respect to the parameters. However, one can only
view one focus point at a time and 
the approach does not scale well to many dimensions.
%either ($O(n^2)$) which means that for 
With each additional dimension one must substantially 
shrink the subplots so less detail can be seen just like in scatterplot matrices. Topological and dimensionality
reduction techniques take the opposite approach. They morph the space to
produce a 2D global view. However, the morphing process is rather
complex and it is unclear what the resulting layout means. This reduces
comprehension. Ideally we would like to somehow combine the global view with
local detail.

In this chapter, I explore the idea of 1D slices to fill this gap. I focus on
multi-dimensional continuous scalar functions. I define these as functions that
take two or more scalar parameters as input and produce a single scalar as
output.  Towards illustrating the benefits, I propose a concrete technique
using projections of 1D slices, called Sliceplorer~\cite{Torsney-Weir:2017a}
(\autoref{fig:walkthrough}).  1D slices are traditional function line plots
familiar to anyone with basic mathematics knowledge: one dimensional curves
with respect to changes in a single parameter.  Like HyperSlice, they show a
separate subplot for each input dimension. For 1D slices, the number of
subplots scales linearly with the number of dimensions, not quadratically as
with 2D slices. I address the issue of having to choose a focus point by
sampling focus points and then showing all slices as a projection. Therefore,
my 1D approach can be seen as a hybrid method of slicing and projection
techniques. 

In order to evaluate any new technique, we need to
consider what data characteristics and tasks each method is good for. As of yet, there has not been a comprehensive
listing of the tasks a user would want to perform when looking at
multi-dimensional continuous functions. To this
end, I begin development of this task summary by extending the task
classification of Amar, Eagan, and Stasko~\cite{Amar:2005}. I use this
classification to evaluate Sliceplorer and compare it to 2D slices and
different topological approaches. This comparison characterizes
which technique is best for which tasks and reveals that 1D slices is the most
flexible of the current approaches. That is, it supports the broadest range of
different tasks.

I also provide three usage scenarios comparing Sliceplorer with other state of the art techniques. These scenarios illustrate that 1D
slices can reveal structure in the functions that could not
have been seen before. For example, I discuss how one can use my
method to compare the \emph{global} prediction manifold of a neural
network algorithm against a support vector machine to guide and better understand the architecture of a neural network. This is currently an open research question in the machine learning community.

In summary my contributions are: 
\begin{itemize}
\item exploring the idea of 1D slices with a concrete software prototype called Sliceplorer,
\item a task-based analysis of multi-dimensional functions, comparing 1D slices to 2D HyperSlices and topological techniques, and
\item %evaluating our technique through 
three usage scenarios illustrating the value of 1D slices in common function 
      visualization scenarios. 
\end{itemize}


