%Thesis: Multi-dimensional spaces are better visualized through slice based views

We live in a three-dimensional world. This is a relatively
uncontroversial statement. Some argue that time also behaves as a
dimension but we do not have control over it thus giving the universe
3.5 dimensions~\cite{cite}. Nevertheless, ourselves and what we can
interact with are in three dimensions.

However, we interact with many phenomena around us that are essentially
multi-dimensional in nature. For example, the weather in a certain
location is determined by the temperature, pressure, humidity, dew
point, wind velocity, wind direction, among others. A change in any of
these factors results in a change in the weather. Each of these factors
can be viewed as a dimension of some space\ttwnote{spatial embedding}.
By ``walking'' or navigating through this space we can observe the
effect on the weather due to changes in these parameters. By exploring
the relationship between input factors (temperature, pressure, et
cetera) and outputs (the weather) we can build an understanding on how
these factors infuence the weather.

Understanding multi-dimensional continuous spaces is difficult. As
three-dimensional beings we have real-world analogs for measurement,
angle, and position in three dimensions. We do not have these once we
move beyond three dimensions though. Nevertheless, visual analysis of
these multi-dimensional spaces has produced insights about the
underlying behavior~\cite{Sedlmair:2014}. The issue is how to show more
than three dimensions on a two-dimensional screen. There are several
techniques to do this: dimension reduction, projection, slicing,
topology, linked views, and slicing.

Slice-based views of multi-dimensional continuous spaces have not been
explored as extensively as other options. 
This work began with HyperSlice~\cite{Wijk:1993}. HyperSlice provides the
framework for visualizing multi-dimensional continuous objects as a set of
two-dimensional slices. \ttwnote{more description}
In my own work, I aim to explore the possibilities of
\ttwnote{I aim for work covering all possible data types}
In this thesis I will explore
the possibilities of these slice-based views. Through a number of case
study examples, I will demonstrate the power of these views and ways to
address their shortcomings.

\input{motivation/multid_spaces}
\input{motivation/challenges}
\input{motivation/advantages}
\input{motivation/slicing_work}
\input{motivation/overview}

